Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex

\bcp{Agreed.  I think in particular the first paragraph can just go.
  But this is not enough.  Also, the section is trying to do at least
  two things---talk about earlier work and how we compare, and talk
  about why we believe a flexible approach is good.  The whole
  discussion is a bit wooly and confusing.}

@@

Security properties should be motivated by security requirements
and flexible enough to support various sets of security requirements depending on the
goals of the system that they validate.
Following this reasoning, our vision of stack safety is as a collection of properties
describing the security needs of a system with regard to its functional abstractions.
While the stack still plays a key role, our stack safety is about functions. The most
dramatic consequence of this is that we treat registers as part of a function activation's
state, and hold them to the same protection requirements as stack memory.

Flexibility is especially important for stack safety.
%
Most security properties are described at the level of a high-level programming language,
and translated to the target-level via a secure compiler. Stack safety cannot
be defined in this way, as the abstraction is completely transparent in most source languages.
(Contrast Azevedo de Amorim et al.'s \cite{DBLP:conf/post/AmorimHP18} work on heap safety,
which is broadly consistent across language semantics, and therefore
amenable to a very general high-level approach.)
%
But it is not satisfying as a low-level property; at the level of the ABI, the
specification of a well-behaved stack is minimal. The ABI is not concerned with such
questions as whether a caller's frame should be readable or writable to its callee,
leaving those concerns to be determined by a high-level language built atop the stack.
These are high-level security questions.

@@

In particular, many enforcement techniques focus purely on
well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
to enforce it completely. Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
When checking if a preliminary version of these properties applied to
the Cerise calling convention, they noted that the Skostengaard et al. calling convention
did not enforce callee confidentiality, and made sure that their version would
\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
This demonstrates the benefit of our choice to explicitly state properties in security
terms: specifying security mechanisms is hard, and when the specification takes the
form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
security requirement.

@@

      ~ \stackrel{\mathtt{a0}}{\objbox\objbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}

>>>>

\bcp{Agreed.  I think in particular the first paragraph can just go
  (I've deleted it).
  But this is not enough.  Also, the section is trying to do at least
  two things---talk about earlier work and how we compare, and talk
  about why we believe a flexible approach is good.  The whole
  discussion is a bit wooly and confusing.}

@@

% Security properties should be motivated by security requirements
% and flexible enough to support various sets of security requirements depending on the
% goals of the system that they validate.
% Following this reasoning, our vision of stack safety is as a collection of properties
% describing the security needs of a system with regard to its functional abstractions.
% While the stack still plays a key role, our stack safety is about functions. The most
% dramatic consequence of this is that we treat registers as part of a function activation's
% state, and hold them to the same protection requirements as stack memory.

Flexibility is important for talking rigorously about stack safety.
%
The reason is that, while many security properties can be described at
the level of a high-level programming language and translated to a
target machine by a secure compiler, stack safety cannot be defined in
this way, since ``the stack'' is completely transparent in the
definitions of most source languages.\footnote{Contrast Azevedo de
  Amorim et al.'s work on heap safety
  \cite{DBLP:conf/post/AmorimHP18}: the concept of the heap figures
  directly in high-level language semantics and its security is
  therefore amenable to a high-level treatment.}
%
But neither is it satisfying as an entirely low-level property;
indeed, at the lowest level, the specification of a ``well-behaved
stack'' is almost vacuous. The ISA is not concerned with such
questions as whether a caller's frame should be readable or writable
to its callee: those are the purview of high-level languages built
atop the hardware stack.

\bcp{There's a broken transition here: the paragraph below is not an
  instance of the paragraph above.  Indeed, I don't see how they
  connect at all.}

@@

In particular, many enforcement techniques focus purely on
well-bracketed control flow. For instance, stack canaries aim to prevent certain attacks on the return
address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
to enforce it completely. Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by
Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
When checking if a preliminary version of these properties applied to
the Cerise calling convention, they noted that the Skostengaard et al. calling convention
did not enforce callee confidentiality, and made sure that their version would
\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
This demonstrates the benefit of our choice to explicitly state properties in security
terms: specifying security mechanisms is hard, and when the specification takes the
form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
security requirement.

@@

      ~ \stackrel{\mathtt{a0}}{\pubbox\pubbox} ~ \stackrel{\mathtt{a1}}{\freebox\freebox}
